{\fontfamily{ptm}\selectfont Relaxed Weak Accelerated Proximal Gradient Method: a Unified Framework for Nesterov's Accelerations }

### 2025-03-03 Notebook LaTEX/R-WAPG/relaxed-wapg-02-03-2025.tex
>w<, changes

% \input{presets/wang_full.tex}

\usepackage{ifthen}\newboolean{draftmode}\setboolean{draftmode}{false}
\input{presets/wang/input.tex}
% \input{presets/misc.tex}
% \input{presets/julia_lstlisting.tex}

@@

\begin{abstract}
    \noindent
    This paper is devoted to the study of accelerated proximal gradient methods where the sequence that controls the momentum term doesn't follow Nesterov's rule.
    We propose a relaxed weak accelerated proximal gradient (R-WAPG) method, a generic algorithm that unifies the convergence results for strongly convex and convex problems where the extrapolation constant is characterized by a sequence that is much weaker than Nesterov's rule.
    Our R-WAPG provides a unified framework for several notable Euclidean variants of FISTA and verifies their convergences.
    In addition, we provide the convergence rate of strongly convex objective with a constant momentum term.
    We also reformulate R-WAPG so that it doesn't require any parameter, without using the idea of restarting, which we call ``Free R-WAPG".
    Explorative numerical experiments were conducted to show its competitive advantages.
\end{abstract}

@@

\section{Introduction}
    \subsection{Motivations}
        The Nesterov's acceleration scheme originally proposed in 1983 \cite{nesterov_method_1983} is a celebrated first order method for solving the minimization problem.
        If the minimizer $x^*$ exists, then the optimality gap $F(x_k) - F(x^*)$ decreases at a rate of $\mathcal O(1/k^2)$, a faster rate compared to gradient descent which is $\mathcal O(1/k)$.
        It's consider optimal in a certain sense, see Chapter 2 of Nesterov's book \cite{nesterov_lectures_2018} for details.
        However, things are not sunshine, rainbow and happily ever after because the above scheme is associated with the issue that if $F$ is $\mu > 0$ strongly convex, the optimality gap $F(x_k) - F(x^*)$ oscillates and converges sub-linearly, making it worse than gradient descent. See page 9 of Su et al. \cite{su_differential_2016}.
        \par
        This issue motivated a vast amount of literatures aims at improving, interpreting, and extending this method.
        Restarting is a popular solution to address the issue of obtaining better convergence rate when the objective function is strongly convex.
        Beck and Teboulle \cite{beck_fast_2009} mitigate the issue by restarting and showed that it has an $\mathcal O(1/k^2)$ convergence still, but it performs empirically better.
        See \cite{aujol_parameter-free_2024}, \cite{necoara_linear_2019} and references therein for recent advancements in restarting accelerated proximal gradient algorithm.
        \par
        In the convex settings, the Variants Nesterov's Accelerated Gradient optimal to for cases when $\mu \ge 0$ and $\mu > 0$ are different.
        In this paper, we will show a unified framework for the convergence proof of both cases with a much relaxed condition for the recurrence of the momentum parameters used to extrapolate the iterates.
        For simplicity consider minimizing a differentiable convex function $F: \RR^n \rightarrow \RR$.
        Let $F$ be $L$-Lipschitz smooth and $\mu \ge 0$ strongly convex meaning that there exists $L \in \RR$ such that for all $x \in \RR^n, y\in \RR^n$ it has $\mu\Vert x - y\Vert\le \Vert \nabla F(x) - \nabla F(y)\Vert \le L \Vert x - y \Vert$.
        We will list the following example algorithm of Nesterov's Accelerated Gradient to refer to while introducing the context and prior works.
        Initialize $x_1 = y_1$ and $t_0 = 1$, the algorithm finds $(x_k)_{k \ge 1}$ for all $k \ge 1$ by:
        \begin{align}
            & x_{k + 1} = y_k - L^{-1}\nabla F(y_k),
            \\
            & t_{k + 1} = 1/2\left(1 + \sqrt{1 + 4t_{k}^2}\right),
            \\
            & \theta_{k + 1} = (t_{k} - 1)/t_{k + 1},
            \\
            & y_{k + 1} = x_{k + 1} + \theta_{k + 1}(x_{k + 1} - x_k).
        \end{align}\label{eqn:example-algorithm}

        \subsection{Prior works}
            Apidopoulos et al. \cite{apidopoulos_convergence_2018} showed that a relaxed choice of the sequence that doesn't strictly follow the Nesterov's rule still leads to the convergence of the momentum algorithm.
            Besides restarting, See Attouch, et al. \cite{attouch_first-order_2022}, Maulen and Peypouquet \cite{maulen_speed_2023} and references within.
            They present the idea of Hessian Damping for stabilizing the Nesterov's accelerated gradient without/with restarting.

@@

        Our contributions are two folds, theoretical and practical.
        Our results are based the assumption $F = f + g$ where $g:\RR^n \rightarrow \overline\RR$ is convex proper and closed, and $f$ is an $L$-Lipschitz smooth and $\mu \ge 0$ strongly convex function.

@@

        Secondly, we present an alternative to restarting that performs well empirically inspired by a detail in the convergence proof of R-WAPG with extensive numerical experiments.
        \textbf{A summary of our main results follow. }
        Nesterov's acceleration extrapolate $y_{k + 1} = x_{k + 1} + \theta_{k + 1}(x_{k + 1} - x_k)$ where $\theta_{k + 1} = (t_{k} - 1)/t_{k + 1} \in (0, 1)$ is the ``momentum''.
        The choices for $\theta_k$ varies for different variants of the accelerated proximal gradient algorithm.
        In Chambolle and Dossal \cite{chambolle_convergence_2015}, it has $t_k = (n + a - 1)/a$ for all $a > 2$ which gives weak convergence of the iterates $x_k$ in Hilbert space.
        In Chapter 10 of Beck's Book \cite{beck_first-order_2017}, a variant called V-FISTA can achieve the faster linear convergence rate: $\mathcal O((1 - \sqrt{\mu/L})^k)$ on the optimality gap for $\mu > 0$ strongly convex $F$.
        V-FISTA has $\theta_t = (\sqrt{\kappa} - 1)/(\sqrt{\kappa} + 1)$ where $\kappa = \mu/L$.
        We relax the traditional choice of the sequence $\theta_k$ in Equation \eqref{eqn:example-algorithm} and showed an upper bound of the optimal gap.

@@

        Our first main result shows that if $\theta_{k + 1} = (\rho_k\alpha_k(1 - \alpha_k)/(\rho_k\alpha_k^2 + \alpha_{k + 1}))$, using the R-WAPG we proposed in Definition \ref{def:wapg} with Proposition \ref{prop:wapg-convergence}, \ref{prop:r-wapg-momentum-repr}, we can show that the gap $F(x_k) - F(x^*)$ is bounded by:

@@

        Our second main result shows that for any $\alpha_k \ge 0$ the choice of sequence $\alpha_k = a/(a + k)$ results in $\rho_k > 1$ for all $k \in \N$ such that R-WAPG reduces to a variant of FISTA proposed in Chambolle and Dossal \cite{chambolle_convergence_2015}, and we are able to show the same convergence rate in Theorem \ref{thm:r-wapg-on-cham-doss}.
        When $\rho_k = 1, \mu = 0$, R-WAPG reduces perfectly to FISTA by Beck \cite{beck_first-order_2017}, if $\mu > 0, \rho_k = 1$, it reduces to the V-FISTA by Beck \cite{beck_first-order_2017}.
        In Theorem \ref{thm:fixed-momentum-fista}, it demonstrates that R-WAPG frameworks gives a linear convergence claim for all fixed momentum method where $\alpha_k := \alpha \in (\mu/L, 1)$ and  $F$ is $\mu > 0$ strongly convex.
        Finally, we did the necessary work and present three equivalent forms of R-WAPG in Section \ref{sec:rwapg-equiv-repr} that are comparable to notable Euclidean variants of FISTA and beyond. This shows the descriptive power of our R-WAPG framework.
        Our practical contribution is an algorithm inspired by a detail in the proof of Proposition \ref{prop:wapg-convergence} which is called ``Parameter Free R-WAPG'' (See Algorithm \ref{alg:free-rwapg}).
        The algorithm is parameter free, meaning that it doesn't require knowing $L, \mu$ in advance, and it determines the value of $\theta_k$ by estimating the local concavity using iterates $y_{k}, y_{k + 1}$ from the Bregman divergence of $f$ with minimal computational cost.
        We conducted ample amount of numerical experiments to show that it has a favorable convergence rate in practice and behaves similarly to the FISTA with monotone restart.

@@

        problem considered in the paper and give a proximal inequality connecting the objective function
         and the proximal gradient

@@

        a key inequality based on a single generic iterative step that acts similar to a descent lemma.
        Section \ref{sec:rwapg-formulation-convergence} formulates the full R-WAPG in Definition \ref{def:wapg} and defines the R-WAPG sequence in Definition \ref{def:rwapg-seq}. Using Proposition \ref{prop:stepwise-lyapunov} from Section~\ref{sec:stepwise-stuff}

@@

        Section \ref{sec:rwapg-equiv-repr} focuses on three equivalent representations of the R-WAPG algorithms that are comparable to instances of APG found in the literatures.
        Section \ref{sec:rwapg-literatures} formulates FISTA, and V-FISTA sequences as instance of the R-WAPG sequences.
        This section proves the convergence rate of several variants of FISTA using the equivalent forms introduced in Section \ref{sec:rwapg-equiv-repr} and the convergence rate developed Section \ref{sec:rwapg-formulation-convergence}.
        Finally, Section \ref{sec:free-rwapg} gives a parameter free version of R-WAPG algorithm and showcase its numerical experiments for regression, LASSO.

@@

    \begin{assumption}[Smooth and nonsmooth additives]\label{ass:smooth-nsmooth-additive}

@@

    The following lemma provides important properties of $\mathcal M^{L^{-1}}(\cdot; y), \widetilde{\mathcal M}^{L^{-1}}(\cdot; y)$.

@@

    \begin{theorem}[Proximal inequality]\label{thm:prox-grad-ineq}

@@

            \text{By Lemma \ref{lemma:pg-envelope}}

@@

\section{Stepwise formulation of weak accelerated proximal gradient}\label{sec:stepwise-stuff}
    The goal of this section is to build the R-WAPG algorithm which is described in the Definition \ref{def:wapg} of the next section.
    \par
    Definition \ref{def:stepwise-wapg} describes one iteration of the R-WAPG algorithm.
    It defines the procedure of generating $x_{k + 1}, v_{k + 1}$ given any $x_k, v_k$.
    Proposition \ref{prop:stepwise-lyapunov} states the inequality \eqref{ineq:stepwise-lya-2}  of a decreasing quantity involving $F(x_k), F(x_{k + 1})$ at each iteration.
    \par
    The assumption below is crucial to describing the convergence rate of the R-WAPG, and it requires the convexity assumption on $F$.
    \begin{assumption}\label{ass:regret-gap}
        Given $x_k, y_k, v_k$ where $k \in \mathbb Z_+$, we define the following quantities
        \begin{align}
            g_k &\defeq L(y_k - T_L(y_k)),
            \label{eqn:grad-map}
            \\
            l_F(x; y_k) &\defeq F(T_L(y_k)) + \langle g_k, x - y_k\rangle + \frac{1}{2L}\Vert g_k\Vert^2,
            \label{eqn:lower-linearization}
            \\
            \epsilon_{k} &\defeq F(x_k) - l_F(x_k; y_k).
            \label{eqn:regret}
        \end{align}
        % To see $\epsilon_k \ge 0$, use Theorem \ref{thm:prox-grad-ineq} and let $y = y_k, x = x_k$, by the definition $g_k = L(y_k - T_L(y_k))$ it has:
        % \begin{align*}
        %     F(x_k) - F(T_L(y_k))
        %     - \langle L(y_k - T_L(y_k)),x_k - y_k \rangle
        %     - \frac{L}{2}\Vert y_k - T_Ly_k\Vert^2
        %     - \frac{\mu}{2}\Vert x_k - y_k\Vert^2
        %     &\ge 0
        %     \\
        %     \iff
        %     F(x_k) - F(T_L(y_k))
        %     - \langle g_k,x_k - y_k \rangle
        %     - \frac{1}{2L}\Vert g_k\Vert^2
        %     &\ge 0.
        % \end{align*}
    \end{assumption}

    \begin{definition}[Stepwise weak accelerated proximal gradient]\label{def:stepwise-wapg}\;\\

@@

        Fix any $k \in \mathbb Z_+$.

@@

            \hat \gamma_{k + 1} &= (1 - \alpha_k)\gamma_k + \mu \alpha_k, \label{eqn:stepwise-wapg-eqn1}

@@

            g_k &= \mathcal G_L (y_k), \label{eqn:stepwise-wapg-eqn3}
            v_{k + 1} &= \hat\gamma^{-1}_{k + 1}

@@

            x_{k + 1} &= T_L(y_k). \label{eqn:stepwise-wapg-eqn5}

@@

        Then for any $x^* \in \RR^n$ it satisfies:

@@

        The proof starts by showing that \eqref{eqn:Q1} and \eqref{eqn:Q2} below 
        follow from \eqref{eqn:stepwise-wapg-eqn2}, then the desire results follows naturally using algebra.
        We make the following observations from Definition \ref{def:stepwise-wapg}:

@@

            &= \alpha_k \gamma_k(v_k - x_k)

@@

@@

    \begin{proposition}[Stepwise Lyapunov]\label{prop:stepwise-lyapunov}\;\\

@@

        Building upon these intermediate results, we are able to claim \eqref{ineq:stepwise-lya-2}.
        It follows by subtracting $F(x^*)$ on \eqref{eqn:stepwise-lya-proof-step-3} which gives:

@@

        \textbf{Proof of \eqref{eqn:stepwise-lya-proof-step-1}}:

@@

        \textbf{Proof of \eqref{eqn:stepwise-lya-proof-step-2}}:

@@

        \textbf{Proof for \eqref{eqn:stepwise-lya-proof-step-3}}:

@@

        The last equality re-arranged terms and grouped the coefficients of $\epsilon_k$ together.

@@

@@

@@

@@

@@

@@

            \\ &=

@@

@@

        \end{split} % \tag{3.1*}

@@

        \eqref{eqn:stepwise-lya-proof-step-3} is now justified.

    \end{proof}
    \begin{remark}
        Given $y_k$, we can choose to increase $\mu_k = 2D_f(x^*, y_k)/\Vert y_k - x^*\Vert^2 \ge \mu$ which also works on \eqref{ineq:stepwise-lya-proof-ineq4}.
        This is true by the intermediate steps taken to prove Theorem \ref{thm:prox-grad-ineq}.
        $\mu$ is a pessimistic choice for the inequality above.
        But in general the choice of $\mu$ remains the strong convexity modulus or equivalently, any value that is smaller than the true strong convexity constant for claiming the convergence rate for all initial guesses.
    \end{remark}

@@

    In this section we propose Relaxed Weak Accelerated Proximal Gradient (R-WAPG), see Definition \ref{def:wapg}.
    R-WAPG algorithm generates iterates $(x_k, y_k, v_k)$ and admits an upper bound on $F(x_k) - F^*$ described in Proposition \ref{prop:stepwise-lyapunov}.
    Definition \ref{def:rwapg-seq} introduces the concept of an R-WAPG sequences: $(\alpha_k)_{k \ge 0}, (\rho_k)_{k \ge 0}$ which is crucial.
    The sequences parameterize the R-WAPG algorithm stated in Definition \ref{def:wapg}, it connects the step-wise formulation of R-WAPG (Definition \ref{def:stepwise-wapg}) and can describe the convergence claim of R-WAPG in Proposition \ref{prop:wapg-convergence}.
    In the next section, it continues to play a crucial role in describing several equivalent forms of the R-WAPG algorithm, and their corresponding convergence claim.

@@

        We call $(\alpha_k)_{k \ge 0}, (\rho_k)_{k \ge 0}$ the \textbf{R-WAPG Sequences}.

@@

    Not all properties of the R-WAPG sequences are used in the convergence proof, some other of its properties such as $\alpha_k \in (\mu/L, 1)$ plays a role much later on in Section \ref{sec:rwapg-equiv-repr}.
    The following lemma states two consequences of the definition of the R-WAPG sequences which will play a crucial role in the convergence proof immediately after.
    \begin{lemma}{Consequences from the definition of R-WAPG Sequences}\label{lemma:r-wapg-seq-consequences}
        Let sequence $(\alpha_k)_{k\ge 0}$ and $(\rho_k)_{k \ge0}$ be given by Definition \ref{def:rwapg-seq}, then they have:

@@

            \item $\rho_k > 0$ for all $k \ge 0$.
            \item $\forall k \ge 1: L\alpha_k^2 = (1 - \alpha_k)L\rho_{k - 1}\alpha_{k - 1}^2 +\mu\alpha_k$.

@@

        The first item is obvious by $\alpha_k \in (\mu/L, 1)$.
        The second item is direct by definition of $\rho_k$ which has for all $k \ge 1$:

@@

    \begin{definition}[Relaxed weak accelerated proximal gradient (R-WAPG)]\label{def:wapg}\;\\

@@

            &= (1 - \alpha_k)\gamma_k + \alpha_k \mu

@@

            &= (1 - \alpha_k)\rho_{k - 1}\hat \gamma_{k} + \alpha_k \mu.
        Therefore, the sequences $(\alpha_k)_{k \ge 0}, (\rho_k)_{k \ge 0}$ fits the same recurrence relation from Lemma \ref{lemma:r-wapg-seq-consequences} (ii).

@@

    \begin{proposition}[R-WAPG convergence claim]\label{prop:wapg-convergence}\; \\

@@

            & \hat \gamma_2 = L \alpha_1^2 , \text{ Definition \ref{def:wapg}}

@@

            & \quad \text{By Proposition \ref{prop:stepwise-lyapunov}}

@@

    The goal of this section is to reduce Definition \ref{def:wapg} into simpler forms that are equivalent to what commonly appears in the literatures.
    In the literatures, variants of Accelerated Proximal Gradient algorithm such as FISTA, V-FISTA has different representations.
    Sometimes the same algorithm include two iterates $(x_k, y_k)$, sometimes they include three iterates $(x_k, y_k, z_k)$.

>>>>

% \input{presets/wang_full.tex}
\usepackage{amsmath}
\allowdisplaybreaks
\usepackage{ifthen}\newboolean{draftmode}\setboolean{draftmode}{false}
\input{presets/wang/input.tex}
% \input{presets/misc.tex}
% \input{presets/julia_lstlisting.tex}

@@

\begin{abstract}
    \noindent
    This paper is devoted to the study of accelerated proximal gradient methods where the sequence that controls the momentum term doesn't follow Nesterov's rule.
    We propose a relaxed weak accelerated proximal gradient (R-WAPG) method, a generic algorithm that unifies the convergence results for strongly convex and convex problems where the extrapolation constant is characterized by a sequence that is much weaker than Nesterov's rule.
    Our R-WAPG provides a unified framework for several notable Euclidean variants of FISTA and verifies their convergences.
    In addition, we provide the convergence rate of strongly convex objective with a constant momentum term.
    Without using the idea of restarting, we also reformulate R-WAPG as ``Free R-WAPG" so that it doesn't require any parameter.
    Explorative numerical experiments were conducted to show its competitive advantages.
\end{abstract}

@@

\section{Introduction}
    \subsection{Motivations}
        The Nesterov acceleration scheme, originally proposed in 1983 \cite{nesterov_method_1983}, is a celebrated first order method for solving the minimization problem.
        If the minimizer $x^*$ exists, then the optimality gap sequence $(F(x_k) - F(x^*))_{k\geq 1}$ decreases at a rate of $\mathcal O(1/k^2)$, a faster rate compared to gradient descent which is $\mathcal O(1/k)$.
        It's consider optimal in a certain sense, see Chapter 2 of Nesterov's book \cite{nesterov_lectures_2018} for details.
        However, things are not sunshine, rainbow and happily ever after because the above scheme is associated with the issue that if $F$ is $\mu > 0$ strongly convex, the optimality gap sequence $(F(x_k) - F(x^*))_{k\geq 1}$ oscillates and converges sub-linearly, making it worse than the gradient descent. See page 9 of Su et al. \cite{su_differential_2016}.
        \par
        This issue motivated a vast amount of literatures aims at improving, interpreting, and extending this method.
        Restarting is a popular solution to address the issue of obtaining better convergence rate when the objective function is strongly convex.
        Beck and Teboulle \cite{beck_fast_2009} mitigate the issue by restarting and showed that it has an $\mathcal O(1/k^2)$ convergence still, but it performs empirically better.
        See \cite{aujol_parameter-free_2024}, \cite{necoara_linear_2019} and references therein for recent advancements in restarting accelerated proximal gradient algorithms.
        \par
        In the convex setting, the variants that Nesterov's Accelerated Gradient are optimal to for cases when $\mu \ge 0$ and $\mu > 0$ are different.
        In this paper, we will show a unified framework for the convergence proof of both cases with a much relaxed condition for the recurrence of the momentum parameters used to extrapolate the iterates.
        For simplicity consider minimizing a differentiable convex function $F: \RR^n \rightarrow \RR$.
        Let $F$ be $L$-Lipschitz smooth and $\mu \ge 0$ strongly convex meaning that there exists $L \in \RR$ such that for all $x \in \RR^n, y\in \RR^n$ it has $\mu\Vert x - y\Vert\le \Vert \nabla F(x) - \nabla F(y)\Vert \le L \Vert x - y \Vert$.
        We will list the following example algorithm of Nesterov's Accelerated Gradient to refer to while introducing the context and prior works.
        Initialize $x_1 = y_1$ and $t_0 = 1$, the algorithm finds $(x_k)_{k \ge 1}$ for all $k \ge 1$ by:
        \begin{align}
            & x_{k + 1} := y_k - L^{-1}\nabla F(y_k),
            \\
            & t_{k + 1} := 1/2\left(1 + \sqrt{1 + 4t_{k}^2}\right),
            \\
            & \theta_{k + 1} := (t_{k} - 1)/t_{k + 1}, \label{eqn:example-algorithm}
            \\
            & y_{k + 1} := x_{k + 1} + \theta_{k + 1}(x_{k + 1} - x_k).
        \end{align}

        \subsection{Prior works}
            In \cite{apidopoulos_convergence_2018} Apidopoulos et al. showed that a relaxed choice of the sequence that doesn't strictly follow Nesterov's rule 
            still leads to the convergence of the momentum algorithm. In \cite{aujol_parameter-free_2024}, Aujol et al. gave a parameter free FISTA by adaptive restart and backtracking. 
           In Attouch, et al. \cite{attouch_first-order_2022}, Maulen and Peypouquet \cite{maulen_speed_2023}. the authors
            present the idea of Hessian Damping for stabilizing Nesterov's accelerated gradient method without/with restarting.

@@

        Our are concerned with the composite model
        $\min_{x\in\RR^n}\{F(x)= f(x) + g(x)\},$ 
        where $g:\RR^n \rightarrow \overline\RR$ is a proper,  closed and convex function, and $f$ is an $L$-Lipschitz smooth and $\mu \ge 0$ strongly convex function.
        Our contributions are two folds: theoretical and practical.

@@

        Secondly, as an alternative to restarting, we present the method of Free R-WAPG which estimates the modulus of convexity as the algorithm executes. Extensive numerical experiments illustrate that our method performs well empirically.
        \textbf{A summary of our main results follows. }
        Nesterov's acceleration extrapolates $y_{k + 1} := x_{k + 1} + \theta_{k + 1}(x_{k + 1} - x_k)$ where $\theta_{k + 1} := (t_{k} - 1)/t_{k + 1} \in (0, 1)$ is the ``momentum''.
        The choices for $\theta_k$ vary for different variants of the accelerated proximal gradient algorithm.
        In Chambolle and Dossal \cite{chambolle_convergence_2015}, it has $t_k := (k + a - 1)/a$ for all $a > 2$ which gives weak convergence of the iterates $x_k$ in Hilbert space.
        In Chapter 10 of Beck's Book \cite{beck_first-order_2017}, a variant called V-FISTA can achieve the faster linear convergence rate: $\mathcal O((1 - \sqrt{\mu/L})^k)$ on the optimality gap for $\mu > 0$ strongly convex $f$.
        V-FISTA has $\theta_k := (\sqrt{\kappa} - 1)/(\sqrt{\kappa} + 1)$ where $\kappa := L/\mu$.
        We relax the traditional choice of the sequence $(\theta_k)_{k\geq 1}$ in equation \eqref{eqn:example-algorithm} and give an upper bound of the optimal gap.

@@

        Our first main result shows that if $\theta_{k + 1} := (\rho_k\alpha_k(1 - \alpha_k)/(\rho_k\alpha_k^2 + \alpha_{k + 1}))$, using the R-WAPG we proposed in Definition \ref{def:wapg} with Proposition \ref{prop:wapg-convergence}, \ref{prop:r-wapg-momentum-repr}, we can show that the gap $F(x_k) - F(x^*)$ is bounded by:

@@

        Our second main result shows that for any $a \ge 2$ the choice of sequence $\alpha_k := a/(a + k)$ results in $\rho_k \geq 1$ for all $k \in \N$ such that R-WAPG reduces to a variant of FISTA proposed in Chambolle and Dossal \cite{chambolle_convergence_2015}, and we are able to show the same convergence rate in Theorem \ref{thm:r-wapg-on-cham-doss}.
        When $\mu := 0, \rho_k := 1$, R-WAPG reduces perfectly to FISTA by Beck \cite{beck_first-order_2017}; when $\mu > 0, \rho_k := 1$, it reduces to the V-FISTA by Beck \cite{beck_first-order_2017}.
        In Theorem \ref{thm:fixed-momentum-fista}, we demonstrate that R-WAPG framework gives a linear convergence claim for all fixed momentum method where $\alpha_k := \alpha \in (\mu/L, 1)$ and  $F$ is $\mu > 0$ strongly convex.
        Finally, we present three equivalent forms of R-WAPG in Section \ref{sec:rwapg-equiv-repr} that are comparable to notable Euclidean variants of FISTA and beyond. This shows the enormous flexibility of our R-WAPG framework.
        Our practical contribution is an algorithm called ``Parameter Free R-WAPG'' (See Algorithm \ref{alg:free-rwapg}), and is is inspired by the proof of Proposition \ref{prop:wapg-convergence}.
        The algorithm is parameter free because it doesn't require knowing $L, \mu$ in advance, and it determines the value of $\theta_k$ by estimating the local concavity using iterates $y_{k}, y_{k + 1}$ from the Bregman divergence of $f$ with minimal computational cost.
        We conduct ample amount of numerical experiments to show that it has a favorable convergence rate in practice and behaves similarly to the FISTA with monotone restart.

@@

        problem considered in the paper and give a proximal inequality connecting the objective function,
        the proximal gradient

@@

        a key inequality based on a single generic iterative step that acts similar to a descent lemma; see Proposition \ref{prop:stepwise-lyapunov}. It is crucial for the convergence analysis of relaxed weak accelerated proximal gradient (R-WAPG) studied later.
        Section \ref{sec:rwapg-formulation-convergence} formulates the full R-WAPG and defines the R-WAPG sequence. Using Proposition \ref{prop:stepwise-lyapunov} from Section~\ref{sec:stepwise-stuff}

@@

        Section \ref{sec:rwapg-equiv-repr} focuses on three equivalent representations of the R-WAPG algorithm that are comparable to instances of APG found in the literatures.
        Section \ref{sec:rwapg-literatures} formulates FISTA, and V-FISTA sequences as the instances of R-WAPG sequences.
        This section proves the convergence rate of several variants of FISTA using the equivalent forms introduced in Section \ref{sec:rwapg-equiv-repr} and the convergence rate developed in Section \ref{sec:rwapg-formulation-convergence}.
        In Section \ref{sec:free-rwapg} we give a parameter free version of R-WAPG algorithm and showcase its numerical experiments for regression, LASSO. We conclude with some discussions and perspectives of our results in Section~\ref{s:discussion}.

@@

    \begin{assumption}[smooth and nonsmooth additives]\label{ass:smooth-nsmooth-additive}

@@

    The following lemma provides important properties of $\mathcal M^{L^{-1}}(\cdot; y)$ and $ \widetilde{\mathcal M}^{L^{-1}}(\cdot; y)$.

@@

    \begin{theorem}[proximal inequality]\label{thm:prox-grad-ineq}

@@

            \text{(By Lemma \ref{lemma:pg-envelope})}

@@

\section{Stepwise formulation of weak accelerated proximal gradient}\label{sec:stepwise-stuff}
    The goal of this section is to provide some preparatory results for the R-WAPG algorithm described in Definition \ref{def:wapg} of the next section.
    Definition \ref{def:stepwise-wapg} below describes one iteration of the R-WAPG algorithm.
    It defines the procedure of generating $x_{k + 1}, v_{k + 1}$ given any $x_k, v_k$.
    The main result of this section is Proposition \ref{prop:stepwise-lyapunov}, which states the key inequality \eqref{ineq:stepwise-lya-2}  of a decreasing quantity involving both $F(x_k)$ and $F(x_{k + 1})$ at each iteration.
    \par
    The assumption below is crucial to describing the convergence rate of the R-WAPG, and it requires the convexity assumption on $F$.
    \begin{assumption}\label{ass:regret-gap}
        Given $x_k, y_k, v_k$ where $k \in \mathbb \NN$, we define the following quantities:
        \begin{align}
            g_k &\defeq L(y_k - T_L(y_k)),
            \label{eqn:grad-map}
            \\
            l_F(x; y_k) &\defeq F(T_L(y_k)) + \langle g_k, x - y_k\rangle + \frac{1}{2L}\Vert g_k\Vert^2,
            \label{eqn:lower-linearization}
            \\
            \epsilon_{k} &\defeq F(x_k) - l_F(x_k; y_k).
            \label{eqn:regret}
        \end{align}
        % To see $\epsilon_k \ge 0$, use Theorem \ref{thm:prox-grad-ineq} and let $y = y_k, x = x_k$, by the definition $g_k = L(y_k - T_L(y_k))$ it has:
        % \begin{align*}
        %     F(x_k) - F(T_L(y_k))
        %     - \langle L(y_k - T_L(y_k)),x_k - y_k \rangle
        %     - \frac{L}{2}\Vert y_k - T_Ly_k\Vert^2
        %     - \frac{\mu}{2}\Vert x_k - y_k\Vert^2
        %     &\ge 0
        %     \\
        %     \iff
        %     F(x_k) - F(T_L(y_k))
        %     - \langle g_k,x_k - y_k \rangle
        %     - \frac{1}{2L}\Vert g_k\Vert^2
        %     &\ge 0.
        % \end{align*}
    \end{assumption}

    \begin{definition}[stepwise weak accelerated proximal gradient]\label{def:stepwise-wapg}\;\\

@@

        Fix any $k \in \NN$.

@@

            \hat \gamma_{k + 1} &:= (1 - \alpha_k)\gamma_k + \mu \alpha_k, \label{eqn:stepwise-wapg-eqn1}
            y_k &:=

@@

            g_k &:= \mathcal G_L (y_k), \label{eqn:stepwise-wapg-eqn3}
            v_{k + 1} &:= \hat\gamma^{-1}_{k + 1}

@@

            x_{k + 1} &:= T_L(y_k). \label{eqn:stepwise-wapg-eqn5}

@@

        Then for any $x^* \in \RR^n$ we have

@@

        We start by showing that \eqref{eqn:Q1} and \eqref{eqn:Q2} below 
        follow from \eqref{eqn:stepwise-wapg-eqn2}, and the desired result can be derived naturally using algebra.
        By Definition \ref{def:stepwise-wapg}:

@@

            = \alpha_k \gamma_k(v_k - x_k)

@@

            = \alpha_k(x^* - x_k).

@@

    Our main result is this section comes as follows.
    \begin{proposition}[stepwise Lyapunov]\label{prop:stepwise-lyapunov}\;\\

@@

        Building upon these intermediate results, we are able to claim \eqref{ineq:stepwise-lya-2}. Indeed,
         subtracting $F(x^*)$ on \eqref{eqn:stepwise-lya-proof-step-3} gives:

@@

        \emph{Proof of \eqref{eqn:stepwise-lya-proof-step-1}}:

@@

        \emph{Proof of \eqref{eqn:stepwise-lya-proof-step-2}}:

@@

        \emph{Proof of \eqref{eqn:stepwise-lya-proof-step-3}}:

@@

        The last equality groups the coefficients of $\epsilon_k$ together.

@@

        %\begin{split}

@@

            \right)\nonumber

@@

                + \frac{\mu \alpha_k}{2}\Vert v_k - x^*\Vert^2 \nonumber

@@

            \nonumber\\

@@

            \nonumber\\ &\quad

@@

            \nonumber\\ &=

@@

            \nonumber

@@

        %\end{split} % \tag{3.1*}

@@

        Hence \eqref{eqn:stepwise-lya-proof-step-3} is justified.
        Given $y_k$, we can choose to increase $\mu_k = 2D_f(x^*, y_k)/\Vert y_k - x^*\Vert^2 \ge \mu$ which also works on \eqref{ineq:stepwise-lya-proof-ineq4}.
        This is true by the intermediate steps taken to prove Theorem \ref{thm:prox-grad-ineq}.
        $\mu$ is a pessimistic choice for the inequality above.
        But in general the choice of $\mu$ remains the strong convexity modulus or equivalently, any value that is smaller than the true strong convexity constant for claiming the convergence rate for all initial guesses.

@@

    In this section we propose \emph{Relaxed Weak Accelerated Proximal Gradient (R-WAPG)}, see Definition \ref{def:wapg}.
    The R-WAPG algorithm generates iterates $(x_k, y_k, v_k)$ and admits an upper bound on $F(x_k) - F^*$ described in Proposition \ref{prop:stepwise-lyapunov}.
    Definition \ref{def:rwapg-seq} introduces the R-WAPG sequences $(\alpha_k)_{k \ge 0}, (\rho_k)_{k \ge 0}$ which are crucial for our analysis.
    These sequences parameterize the R-WAPG algorithm, connect the step-wise formulation of R-WAPG (Definition \ref{def:stepwise-wapg}) and can guarantee the convergence of R-WAPG in Proposition \ref{prop:wapg-convergence}.
    In the next section, they continue to play an 
    essential role in describing several equivalent forms of the R-WAPG algorithm, and their corresponding convergence claims.

@@

        We call $(\alpha_k)_{k \ge 0}, (\rho_k)_{k \ge 0}$ as the {R-WAPG sequences}.

@@

    Not all properties of the R-WAPG sequences are used in the convergence proof, some other properties such as $\alpha_k \in (\mu/L, 1)$ plays a role much later on in Section \ref{sec:rwapg-equiv-repr}.
    The following lemma states two consequences of the R-WAPG sequences, which will play a crucial role in the convergence proof immediately after.
    \begin{lemma}\label{lemma:r-wapg-seq-consequences}
        Let sequence $(\alpha_k)_{k\ge 0}$ and $(\rho_k)_{k \ge0}$ be given by Definition \ref{def:rwapg-seq}. Then 

@@

            \item\label{i:rho} $\rho_k > 0$ for all $k \ge 0$.
            \item\label{i:alpha} $\forall k \ge 1: L\alpha_k^2 = (1 - \alpha_k)L\rho_{k - 1}\alpha_{k - 1}^2 +\mu\alpha_k$.

@@

        \ref{i:rho} is obvious by $\alpha_k \in (\mu/L, 1)$.
        \ref{i:alpha} is direct by definition of $\rho_k$ which has for all $k \ge 1$:

@@

    \begin{definition}[relaxed weak accelerated proximal gradient (R-WAPG)]\label{def:wapg}\;\\

@@

            = (1 - \alpha_k)\gamma_k + \alpha_k \mu

@@

            = (1 - \alpha_k)\rho_{k - 1}\hat \gamma_{k} + \alpha_k \mu.
        Therefore, the sequences $(\alpha_k)_{k \ge 0}, (\rho_k)_{k \ge 0}$ fits the ecurrence relation from Lemma \ref{lemma:r-wapg-seq-consequences}\ref{i:alpha}.

@@

    \begin{proposition}[R-WAPG convergence]\label{prop:wapg-convergence}\; \\

@@

            & (\hat \gamma_2 = L \alpha_1^2 , \text{ Definition \ref{def:wapg}})

@@

            & \quad \text{(By Proposition \ref{prop:stepwise-lyapunov})}

@@

    The goal of this section is to reduce Definition \ref{def:wapg} into various forms that are equivalent to what commonly appear in the literature.
    Variants of Accelerated Proximal Gradient algorithm such as FISTA, V-FISTA have different representations.
    Sometimes the same algorithm include two iterates $(x_k, y_k)$; sometimes they include three iterates $(x_k, y_k, z_k)$.
    In this section we simplify R-WAPG to these representations commonly found in the literature and showed that they are equivalent.
